---
title: "Launching Pet Box Subscriptions"
author: "Grace Hyo-Eun Park"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage[labelformat=empty]{caption}
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage{booktabs}
output: 
  pdf_document:
    latex_engine: xelatex
---

\newpage 
\tableofcontents 
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE,  message = FALSE,
                      fig.align = 'center')
library(tidyverse)
library(knitr)
library(kableExtra)

### Loading some datasets ###
pet_df <- read.csv("data/pet_sales_cleaned.csv")
pet_count <- read.csv("data/pet_product_count.csv")
```

# I. About the Company

**PetMind** is a pet product retailer since 2017 in the United States. The company sells products such as accessory, bedding, clothes, food, equipment, grooming, housing, medicine, snack, supplements, and toys for dogs, cats, fish, birds, hamsters, and rabbits. Currently, their goal is to reduce the cost of customer retention by improving brand loyalty. For the last decade, the subscription economy has grown by more than 435% (Subscription Economy Index, 2022). Hence, their first approach to this is to launch a monthly pet box subscription within the next 2 months by analyzing historical data of PetMind from last year.

# II. Business Task

**Preparing a list of popular products for the pet box subscription.**

The current objective of **PetMind** is to improve brand loyalty by launching a monthly pet box subscription. Hence, the business task is to narrow down the list of popular products that should be included in the monthly subscription. 

In particular, the chief marketing officer, or the stakeholder, is interested in determining whether repurchased products should be the only ones included for the pet box subscription. Two of the major aspects we are interested in investigating for designing subscriptions are (1) repurchase and (2) sales. Also, it is important to note that the pet box subscription should be designed differently for different types of pets.

## II-i. Questions to Consider

1. How many products are being purchased more than once?
2. Do the products being purchased again have better sales than others?
3. What products are more likely to be purchased again for different types of pets?

## II-ii. Why Do People Subscribe?

According to *Pivotal* (2022), people subscribe for the following reasons:

- Convenience: Ability to repurchase through easy and automatic payment for every designated period (e.g., monthly, annually).
- Product tangibility: People feel a stronger connection and attraction to products they can physically receive at their door compared to other services like a video subscription. 
- Adequate balance of control and choice: Subscribing allows one to access a decent amount of options (choice) as well as the ability to pause or resume their subscription (controlling their choice).
- A sense of community: Humans are social animals. Having the feel of connection to a specific brand will make them feel valued as a customer.

\newpage

# III. Process of Data Validation

First, we created a database and table in MySQL Workbench and imported the `pet_sales.csv` file, which added 879 observations to the system. Product ID, Product Category, Sales, Vendor ID, Pet Size, Pet Type, and Re-buy were imported as characters and the others were imported as numeric. Based on the data dictionary, we went over each of the features to verify their integrity and whether the data type is accurate for data analysis. 

First of all, sales of products from last year should be numeric according to the data dictionary (and obviously due to its nature); hence, we converted it to a numeric type after removing the dollar signs ($) and commas (,). While browsing through the types of pets available in the dataset, we noticed there are six different types in total, which includes the following: Birds, Cats, Dogs, Fish, Hamsters, and Rabbits. Since the analysis requires only the first four, 46 rows of hamsters and rabbits were omitted, leaving 833 rows. 

Other variables seemed to be accurate, and this was easily verified with aggregate functions in MySQL such as `COUNT()`, `MAX()`, and `MIN()`, and queries like `GROUP BY`. Product and Vendor ID's are the primary keys since the count of these ID's match the total number of observations in the dataset. The dataset includes exactly 11 different product categories: 137 equipment, 137 toys, 71 supplements, 42 bedding, 72 medicine, 42 housing, 71 food, 42 clothes, 42 accessory, and 42 grooming products were entered in the dataset. Similarly, there are 5 different pet sizes: 98 extra small, 271 small, 270 medium, 97 large, and 97 extra large. Moreover, the ratings fall within the range of 1 to 10 as desired, with approximately 6.53 being the average rating. Finally, the number of repurchased products last year were 390 whereas 443 were not repurchased.

## III-i. Monthly Cost for Each Pet Type

Notice that the dataset does not include the average monthly expenditure for each pet type. Using center values such as mean or median would not be sufficient since they do not capture the average monthly budget of a pet owner. Hence, based on internet research, we found that different types of pets have the following range or average of expenditure (in US dollars):

- Birds: 40 - 50
- Cats: 150 - 160
- Dogs: 150 - 350
- Fish: 40 - 50

Note that this depends on how old the pets are, how long the pet owner has had the pet for, the lifestyle of the pet owners, and many other factors. Since pet box subscriptions are meant to be budget-friendly, we are going to make the cost of pet box subscriptions for each pet type less than the lower range of monthly cost stated above. Therefore, we are going to filter out bird and fish products over \$40 and cat and dog products over \$150. This will result in 427 rows.

```{r manipulating_data}
pet_df <- pet_df %>% mutate(price_scaled = price - mean(price),
                            rating_scaled = rating - mean(rating),
                            sales_scaled = sales - mean(sales),
                            number_sold_yr = round(sales/price, 0),
                            number_sold_m = round((sales/price)/12, 0))

pet_df_filtered <- pet_df %>% 
  filter((pet_type == 'bird' & price <= 40) |
           (pet_type == 'cat' & price <= 150)|
           (pet_type == 'dog' & price <= 150) |
           (pet_type == 'fish' & price <= 40))

write.csv(pet_df_filtered, "data/pet_df_filtered.csv")
```

Now that the data validation is complete, we can move forward to data analysis.

\newpage

# IV. Data Discovery and Visualization

## IV-i. Number of Products being Repurchased

```{r table1_count_rebuy}
table1 <- read.csv("data/table1.csv")
table1 %>% kable(col.names = c("", "Count of Repurchase", "Sum of Sales"),
                 caption = 
                   "Table 1: Count and Sales of Products",
                 booktabs = T) %>% 
  kable_styling(latex_options = "HOLD_position")
```

First, we can create a simple pivot table via Microsoft Excel to find the number of products that were repurchased by customers. 

According to Table 1 above, there are 235 products (`r round(100*235/427,2)`%) that did not get purchased again whereas 192 products (`r round(100*192/427,2)`%) were purchased more than once. Similarly, the sum of sales last year for products that were not repurchased (27,590,000 USD) were greater than those that were repurchased (21,826,000 USD). 

In general, majority of the products were not bought more than once and these kinds of products had higher sales last year compared to repurchased products.

## IV-ii. Number of Repurchased vs. Non-Repurchased Products for each Product Category

Recall that the pet box subscription has to be tailored for different types of pets. Therefore, we need to first identify the popular product categories that are purchased more than once for different types of pets.

```{r fig1}
fig1 <- pet_count %>%
  ggplot(aes(x = product_category, y = count, 
                         fill = as.factor(re_buy))) + 
  geom_bar(position = 'dodge', stat = 'identity') +
  scale_fill_brewer(palette = "Set1", 
                    name = "Repurchased",
                    labels = c("No", "Yes")) +
  labs(x = "Pet type", y = "Count", 
       caption = "Figure 1 \n Number of products that were/were not repurchased \n for different product categories, by different types of pets.") +
  facet_wrap(~pet_type, labeller = as_labeller(c(`bird` = "Bird",
                                                 `cat` = "Cat",
                                                 `dog` = "Dog",
                                                 `fish` = "Fish"))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.75, hjust = 1))

fig1
```

Considering the aggregated data from Table 1, we can visualize the number of product categories that were repurchased for different types of pets (See Figure 1). 

One prominent aspect we observe is that dogs and cats have more products compared to fish and birds. Thus, we might want to consider including more products for dogs and cats compared to fish and birds.

The product categories listed below for each type of pet are those with a higher or equal number of repurchased products than non-repurchased products. Note that the product categories with an asterisk (*) are the ones that have the highest counts among the repurchased products.

- Birds: Equipment, Food, Toys*
- Cats: Clothes, Equipment*, Food, Housing
- Dogs: Bedding, Clothes, Equipment*, Food, Medicine
- Fish: Medicine, Snack, Supplements*, Toys

## IV-iii. Sales of Repurchased vs. Non-Repurchased Products for each Product Category

The sum of sales of each product category last year (See Figure 2) seems to have some overlaps with the findings from Figure 1.

```{r fig2}
fig2 <- pet_df_filtered %>% group_by(pet_type, product_category, re_buy) %>% 
  summarize(sum_sales = sum(sales)) %>% 
  ggplot(aes(x = product_category, sum_sales, fill = as.factor(re_buy))) +
  geom_bar(stat= 'identity', position = 'dodge') +
  facet_wrap(~pet_type, labeller = as_labeller(c(`bird` = "Bird",
                                                 `cat` = "Cat",
                                                 `dog` = "Dog",
                                                 `fish` = "Fish"))) +
  scale_fill_brewer(name = "Repurchased", labels = c("No", "Yes"), palette = "Set1") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.75, hjust = 1)) +
  labs(x = "Product Category", y = "Sum of Sales from Last Year ($)",
       caption = "Figure 2 \n Sales of products that were/were not repurchased \n for different product categories, by different types of pets.")

fig2
```

Based on Figure 2 above, the product categories with higher total sales in repurchased products compared to non-repurchased products are listed below for each type of pet:

- Birds: Equipment*, Food
- Cats: Accessory, Bedding, Clothes, Equipment*, Food, Housing
- Dogs: Bedding, Equipment*, Medicine
- Fish: Medicine*, Supplements

When solely looking at the repurchased products, we have the following listed as the top 5 product categories, in descending order:

```{r table2_sales}
table2 <- read.csv("data/table2.csv")
table2 %>% mutate(pet_type = rep("", nrow(table2))) %>% 
  select(-c(rank)) %>% 
  kable(col.names = c("Pet type", "Product Category", "Total Sales"),
        caption = "Table 2: Top 5 Repurchased Product Categories with High Sales for each Pet Type",
        booktabs = T) %>% 
  pack_rows("Bird", 1, 4) %>% 
  pack_rows("Cat", 5, 9) %>% 
  pack_rows("Dog", 10, 14) %>% 
  pack_rows("Fish", 15, 18) %>% 
  kable_styling(latex_options = "HOLD_position")
```

Based on Table 2, repurchased snacks seem to have relatively high sales for all types of pets. Moreover, Table 2 and Figure 2 suggest that equipment has the highest sales for birds, cats, and dogs.

In general, it is not necessarily the case that repurchased products have higher sales than non-repurchased products. For all pet types, non-repurchased products have higher sales and counts than those of repurchased products (See Table 3 below). Instead, it is the product category that matters within each type of pet since the sales and counts of repurchased products differ for each product category of pet types.

```{r table3_sales_and_rebuy}
table3 <- read.csv("data/table3.csv", row.names = NULL) %>% 
  mutate(X = case_when(
    X == "Sum of sales" ~ "Sum of sales",
    X == "Count of re_buy" ~ "Count of repurchase",
    X == "Total Sum of sales" ~ "Total Sum of sales",
    X == "Total Count of re_buy" ~ "Total Count of repurchase",
    TRUE ~ " "
  ))
table3 <- table3[-c(1, 2, 5, 8, 11),]
row.names(table3) <- NULL
table3 <- rename(table3, c('pet_type' = X, 
                 'No' = Column.Labels, 
                 "Yes" = X.1, 
                 'grand_total' = X.2))
table3 %>% kable(col.names = c("Aggregated data", "No", "Yes", "Grand Total"),
                 caption = "Table 3: Count and Sales of Products for each Pet Type", booktabs = T) %>% 
  add_header_above(c("Pet type" = 1, "Repurchased" = 2, " ")) %>% 
  pack_rows("Bird", 1, 2) %>%
  pack_rows("Cat", 3, 4) %>% 
  pack_rows("Dog", 5, 6) %>% 
  pack_rows("Fish", 7, 8) %>% 
  pack_rows("All pets", 9, 10) %>% 
  kable_styling(latex_options = "HOLD_position")
```

### IV-iii.i. Combining Sales and Repurchase

The repurchase and sales results from above suggest that the following product categories are significant for each type of pet. 

- Birds: Equipment, Food, Snack
- Cats: Equipment, Clothes, Food, Housing, Snack
- Dogs: Bedding, Equipment, Medicine, Snack
- Fish: Medicine, Supplements, Snack

### IV-iv. Relationships between Variables

We can also fit logistic models to investigate whether price, rating, and sales are correlated with repurchase for each type of pet. Results of the logistic models would indicate the log-odds ratios of predictors (for repurchased and non-repurchased products) and their corresponding p-values. The log-odds ratios can be converted to odds ratios by simply exponentiating them. 

```{r table4_mods_results}
bird_df <- pet_df %>% filter(pet_type == 'bird')
cat_df <- pet_df %>% filter(pet_type == 'cat')
dog_df <- pet_df %>% filter(pet_type == 'dog')
fish_df <- pet_df %>% filter(pet_type == 'fish')

bird_mod <- glm(as.factor(re_buy) ~ price_scaled + rating_scaled + 
                  sales_scaled +
                  pet_size, 
             data = bird_df,
           family = binomial(link='logit'))
cat_mod <- glm(as.factor(re_buy) ~ price_scaled + rating_scaled + 
                 sales_scaled + 
                 pet_size, 
             data = cat_df,
           family = binomial(link='logit'))
dog_mod <- glm(as.factor(re_buy) ~ price_scaled + rating_scaled + sales_scaled +
                 pet_size, 
             data = dog_df,
           family = binomial(link='logit'))
fish_mod <- glm(as.factor(re_buy) ~ price_scaled + rating_scaled + 
                  sales_scaled +
                  pet_size, 
             data = fish_df,
           family = binomial(link='logit'))


intercept = c(0.48, 1.04, 1.18, 1.92)
price_scaled = c("1.00", "1.00", "1.00", "1.00")
rating_scaled = c(0.91, 0.97, 0.86, 1.82)
sales_scaled = c("1.00", "1.00", "1.00", "1.00")
pet_size_xs = c(0.46, 0.78, 1.27, "0.00")
pet_size_l = c(1.56, 0.89, 0.69, 0.06)
pet_size_m = c(1.28, 0.88, 1.24, 0.93)
pet_size_s = c(2.85, 0.85, 0.68, 0.38)
pvals <- c(0.47, 0.58, 0.53, 0.88, 0.61, 0.75, 0.81, 0.43,
           "0.90", 0.18, 0.44, 0.91, 0.56, 0.79, 0.71, "0.70",
           0.61, 0.14, "<0.001", "0.90", 0.59, 0.39, 0.62, 0.28,
           0.54, 0.89, 0.06, 0.21, 0.99, 0.12, 0.96, 0.38)

birds_OR <- c(intercept[1], price_scaled[1], rating_scaled[1], 
                   sales_scaled[1], pet_size_xs[1], pet_size_l[1],
                   pet_size_m[1], pet_size_s[1])
cats_OR <- c(intercept[2], price_scaled[2], rating_scaled[2], 
                   sales_scaled[2], pet_size_xs[2], pet_size_l[2],
                   pet_size_m[2], pet_size_s[2])
dogs_OR <- c(intercept[3], price_scaled[3], rating_scaled[3], 
                   sales_scaled[3], pet_size_xs[3], pet_size_l[3],
                   pet_size_m[3], pet_size_s[3])
fish_OR <- c(intercept[4], price_scaled[4], rating_scaled[4], 
                   sales_scaled[4], pet_size_xs[4], pet_size_l[4],
                   pet_size_m[4], pet_size_s[4])

# Create table
mod_results <- data.frame(predictors = 
                            c(rep(c("Intercept", 'Price scaled', 'Rating scaled',
                                         'Sales scaled', 'Pet size - extra small',
                                         'Pet size - large', 'Pet size - medium',
                                         'Pet size - small'), 4)),
                          odds_ratio = c(birds_OR, cats_OR, dogs_OR, fish_OR),
                          p_values = pvals)

mod_results %>% kable(col.names = c("Predictors", "Odds ratio", "P-values"), 
                      caption = "Table 4: Logistic Regression Results",
                      booktabs = T) %>% 
  pack_rows("Bird", 1, 8) %>% 
  pack_rows("Cat", 9, 16) %>% 
  pack_rows("Dog", 17, 24) %>% 
  pack_rows("Fish", 25, 32) %>% 
  kable_styling(latex_options = "HOLD_position")
```

According to the results (See Table 4), price, rating, and sales are not statistically significant for birds, cats, and fish based on the significance level $\alpha = 0.05$, meaning they are not meaningfully correlated with repurchase. A predictor is only statistically significant when it is smaller than $\alpha$^[Although p-value is not the absolute standard all the time; Going in depth and discussing statistics would not be necessary for this report.]. When it comes to dogs, we notice that the p-value of the odds ratio of rating is smaller than $\alpha = 0.05$. This means that the following interpretation is statistically significant:

- For every 1 rating increase in a dog product, a customer repurchasing that product decreases by `r round(1 - exp(summary(dog_mod)$coef[3,1]), 4)*100`% compared to other dog products, given that all other conditions remain constant.

Hence, we might want to consider a dog product with a lower rating when selecting products for the dog pet box subscription. Other factors are not statistically significant for dog products as well.

## IV-v. Summary Statistics of Price

Although product price is not statistically correlated with repurchase, we will still use price to select the products for pet box subscription. This is to ensure that we are not forcing the customers to pay high subscription fees for expensive products within their pet box subscription. Hence, we can visualize the range of prices for each product category of pet type with boxplots. The prices of products we select for box subscriptions would not go beyond the 75th percentile, which is equivalent to the upper end of a box in the plot without the whiskers.

```{r fig3_boxplot}
pet_df_filtered %>% 
  ggplot(aes(x = product_category, y = price, fill = product_category)) +
  geom_boxplot() + 
  facet_wrap(~pet_type, labeller = as_labeller(c(`bird` = "Bird",
                                                 `cat` = "Cat",
                                                 `dog` = "Dog",
                                                 `fish` = "Fish"))) + 
  scale_fill_brewer(palette = "Spectral") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.75, hjust = 1),
        legend.position = "none") +
  labs(x = "Product Category", y = "Price",
       caption = "Figure 3
                  Range of product price for each product category of pet types.") 
```

\newpage 

# V. Recommendations for Monthly Subscriptions

So far, we acquired the list of significant product categories for different types of pets based on repurchase and sales. We also noted that the products for birds and fish should be less than \$40, whereas those for cats and dogs should be less than \$150. Moreover, cats and dogs should have more products in their box subscription due to the fact that they have more products available and the difference in the amount of monthly spend. Rating was also an important factor for dog products. Considering such factors, Table 5 is a list of repurchased products with high sales in the product categories that we have defined as significant. 

```{r table5_specific_products}
table5 <- read.csv("data/table5.csv")

table5 %>% select(-c(pet_type, re_buy)) %>% 
  kable(col.names = c("Pet type, Product Category", "Product ID",
                                  "Price", "Rating", "Sales"),
                    caption = "Table 5: List of Products",
        booktabs = T) %>% 
  pack_rows("Bird", 1, 5) %>% 
  pack_rows("Cat", 6, 15) %>% 
  pack_rows("Dog", 16, 25) %>% 
  pack_rows("Fish", 26, 30) %>% 
  kable_styling(latex_options = "HOLD_position")
```

We will create the pet box subscription as diverse as possible with products from various product categories to make the tangible products more entertaining for customers. Recall that snacks had one of the highest sales among repurchased products (See Table 2). For all pet types, we will also have customers choose a snack sample among a few options given. This would make the subscription feel more unique for themselves and make them feel respected.

The product ID's for each type of pet listed below are the specific products for the corresponding pet box subscription. After a certain amount of time, the products will change to different products within the same product categories to ensure customers would not lose interest.

A discount will be applied to all products in the pet box subscriptions so that the price of products would be cheaper than when they are bought individually. The amount of discount is to be discussed. The prices indicated below are before discounts are applied and before adding a snack sample.

## V-i. Pet Box Subscription - Birds

*Pet Box Price*: \$`r 0.35 + 14.42`

- Equipment: 5037
- Food: 4242
- Snacks: Choose one from 4987, 5005, 5017

## V-ii. Pet Box Subscription - Cats

*Pet Box Price*: \$`r 22.16 + 43.90 + 24.10`

Food: 4583
Clothes: 4663
Equipment: 4797
Snack: Choose one from 4777, 4819, 4639

## V-iii. Pet Box Subscription - Dogs

*Pet Box Price*: \$`r 4.03 + 52.82 + 13.51`

Snack: Choose one from 4430, 4948, 4792, 4816
Equipment: 4932 
Medicine: 4414 
Bedding: 4620

## V-iv. Pet Box Subscription - Fish

*Pet Box Price*: \$`r 22.01 + 36.11`

Snack: Choose one from 5008, 5014
Medicine: 4271
Supplements: 4217 

\newpage 

# VI. Limitations

## VI-i. Monthly Expenditure Per Customer

It would have been more ideal to know the average or median budget of a PetMind customer for each type of pet. That way, it would have been easier to decide on the price for the pet box subscriptions and their corresponding products.

## VI-ii. Survey

As we have seen in the results of logistics regression models for each pet type, the price, rating, or pet size did not significantly correlate with repurchase (except for rating of dogs products). Therefore, it was difficult to identify the factors that led to repurchasing a product. Following-up with the customers with a survey and extracting qualitative or quantitative data could give a better insight in the preferences of customers.

\newpage 

# VII. References

|  Bedford, Emma. "Cats Annual Cost of Care Canada 2021." Statista, 17 Aug. 2021, 
|    https://www.statista.com/statistics/448208/annual-cost-of-caring-for-a-cat-by-type-canada/. 
|  
|  Burgado, Dali. "How Much Does It Cost to Own a Pet?" GEICO Living, GEICO Living, 
|    6 Apr. 2021, 
|    https://www.geico.com/living/home/pets/how-much-does-your-pet-cost-you-per-year/. 
|  
|  Clifford, Garth C. "How Much Does a Parakeet Cost? (+Monthly Expenses)."
|    World Birds, 4 Aug. 2021, 
|    https://worldbirds.com/how-much-does-a-parakeet-cost/. 
|  
|  Patel, Arti. "How Much Does It Cost to Own a Dog? 7 Canadians Break down
|    Their Budgets - National." Global News, Global News, 7 Mar. 2019,
|    https://globalnews.ca/news/4979041/how-much-does-cost-own-dog/. 
|  
|  "The Psychology of Subscription: Why Do We Subscribe?" Pivotal, 16 Mar. 2022,
|    https://www.drivepivotal.com/blog/the-psychology-of-subscription-why-do-we-subscribe/. 
|  
|  "Subscription Economy Index™ vs S&amp;P 500 &amp; U.S. Retail Sales." Zuora,
|    Feb. 2022, https://www.zuora.com/resources/subscription-economy-index/#mktoForm_2180. 